Modeling ice-shelf cavities using Lagrangian elements

### 2018-05-27 Lagrangian_MISOMIP.tex
Updating Lagrangian_MISOMIP

% and freshwater fluxes from icebergs affect regional ocean circulation, sea ice distributions and bottom water formation around Antarctica
\end{abstract}
%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%
% The body of the article must start with a \begin{article} command
%
\clearpage
\vspace{10mm}
\textbf{Key Points:}\\
\begin{itemize}
\item A novel modeling framework is developed to model breakable ice shelves and ice-shelf cavities using Lagrangian elements, held together by numerical bonds.
\item The ocean circulation beneath a (static) Lagrangian ice shelf is almost indistinguishable from the circulation beneath an Eulerian ice-shelf model run in an identical configuration, which  provides a proof of concept for the Lagrangian model. 
\item Enhanced ocean mixing occurs near the ice front after a large calving event, which warms the ocean surface and cools the water column at depth. The cool sub-surface anomalies enter the ice shelf cavity and cause reduced basal melt in the period following the calving event. 
\end{itemize}

@@

At the end of an ice-shelf model time step, ice-shelf fields are aggregated from the elements back onto the Eulerian ocean grid, and are then passed from the ice-shelf model to the ocean model. 
The aggregation from the Lagrangian elements onto the Eulerian grid is done by calculating the fraction of each element's volume that lies in each ocean grid cell, and dividing the fields in proportion to this fraction. For example, the amount of ice mass aggregated onto a given ocean grid cell is found by summing up the masses of all elements which intersect that grid cell, only counting the part of an element's mass that  actually intersects the ocean grid cell. When calculating the intersection between an element and a grid cell, we assume that the elements have surface areas that are shaped as regular hexagons. Seven fields are passed from the ice-shelf models to ocean model: iceberg mass and surface area (used the upper-ocean pressure field), temperature flux, salinity flux and mass flux, and meridian and zonal velocity (used to calculate the momentum flux).

@@

 \section{Experiment Setup}
% the proof of concept for
%\sout{The KID model is used to represent the ice shelf in an idealized ice-shelf cavity MOM6 ocean flow. The simulations are repeated using an Eulerian ice shelf with an identical configuration. The details of the experimental setup for both the Lagrangian and Eulerian models are presented here.}
%
\subsection{Domain configuration}
In order for our simulations to be easily comparable to previous models of ice-shelf cavities, we use an experimental setup based on the configuration created for the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP) \citep{Asay-Davis2016}.
The configuration consists of an idealized ice shelf in a rectangular domain $L_{x}=80$km long and $L_{y}=480$km wide. The ice shelf is grounded on the southern side of the domain with the ice-shelf front at y=650km. The ice thickness and bottom topography of this setup are shown in Figure \ref{fig:ISOMIP_mass_and_topog}. The configuration is the same as that of the Ocean0 setup in the MISOMIP, with three changes made:
\begin{enumerate}
\item The `calving criteria' used in the MISOMIP study (which states that all points in the ice shelf with thickness less than 100m are set to zero thickness) has not been used.
\item The ice shelf has been thickened on the flanks of the domain, so that the latitude of the grounding line increases away from the center of the ice shelf.
\item The ice shelf is configured to be symmetric about its meridional center line ($x=\frac{L_{x}}{2}$). This was achieved by using the average of the left and right flanks of the ice-shelf thickness. 
\end{enumerate}
These three changes were made in order to make the circulation beneath the ice shelf easier to interpret.

\subsection{Ocean Model}
The Lagrangian and Eulerian ice shelves are coupled to the MOM6 ocean model \citep{Hallberg2013}. 
The ocean model is run using a hybrid vertical coordinate system which blends a sigma-level and a z-level coordinate \citep{Stern2017}, implemented using the ALE method \citep{White2009}. In this vertical coordinate, model layers bend underneath surface topography (i.e.: the ice shelf), as they would in a sigma coordinate model, and intersect the bottom topography, as they would in a z-coordinate model. The model has 72 vertical layers and has a horizontal resolution of $\Delta x= 2$ km. The numerical simulations were all repeated using an isopycnal coordinate (without ALE regrinding-remapping). The results were qualitatively similar to the hybrid-coordinate results, and are therefore not presented here.
% All results presented below are from the ALE simulations, unless specified otherwise.

@@

\subsection{Initializing Lagrangian elements:}
%\subsection{Lagrangian and Eulerian ice shelf simulations:}
The Lagrangian ice shelf is created using 10882 Lagrangian hexagonal elements with sides of length $S=0.98$ km . The positions of the hexagonal elements are initialized by packing them together in a space-filling staggered lattice. Gaps along the boundaries are filled in using smaller elements so that the total ice-shelf area is preserved. The initial mass of the ice elements is determined using bilinear interpolation from a prescribed gridded ice mass field.

\subsection{Eulerian ice-shelf:}
We validate the Lagrangian ice shelf model by comparing the results to the results found using an Eularian ice shelf model in an identical static ice shelf configuration (see below).
The Eulerian ice-shelf simulation is performed using an existing Eulerian ice-shelf cavity model \citep{Goldberg2012}, which is an optional module of the the MOM6 ocean model. The ice shelf is initialized on the same grid as the ocean model with a horizontal resolution of $\Delta x = 2$ km. The ice-shelf thickness field is initialized using the same ice-shelf draft used for the Lagrangian model (Figure \ref{fig:ISOMIP_mass_and_topog}). The melt rates in the Eulerian ice-shelf simulation are calculated using the 3 equation model for ice-shelf decay \citep{Holland1999}, with the same parameters used to calculate the ice shelf melt rates in the Lagrangian model.

\subsection{Static ice shelf simulation:}
We test the Lagrangian ice shelf model by simulating the flow beneath the idealized ice shelf with the ice shelf being held statice (i.e.: elements are held stationary). In this simulation the ice shelf is thermodynamically active and is able to `melt' but has a time-invariant thickness (as specified in the Ocean0 experiment in the MISOMIP \citep{Asay-Davis2016}). In this setup, ice-shelf melting generates temperature and salinity fluxes into the ocean, but does not change the thickness of the ice shelf / ice elements. %This can be thought to represent an ice shelf in dynamic equilibrium where the melt is exactly balanced ice-shelf advection. 
A constant wind stress $\vec{\tau} = <\tau_{x},\tau_{y}> = < 0.05, 0.05 >$  $\frac{N}{m^{2}}$ is applied to the ocean surface. Note that this wind stress is was not applied in the original MISOMIP experiments \citep{Asay-Davis2016}.
The model is spun up for 5 years, and the analysis is done on the second 5 years after the spinup.

\subsection{Iceberg calving simulation:}
In a second experiment, the Lagrangian ice shelf model is used to simulate a tabular iceberg detaching from the ice shelf, and drifting into the open ocean. The iceberg calving simulation is initialized using the final state of the static Lagrangian ice shelf simulation (i.e.: at time $t = 10 $ years). When initializing the calving event, we bypass the question of how to prescribe a physical calving law \citep{Benn2007, Alley2008, Levermann2012, Bassis2013} by manually breaking off a semi-circular iceberg.

In the iceberg calving simulation, all ice elements initially within a 14.4~km radius of the center of the ice front to move freely while the other ice elements continue to be held stationary. Ice elements less than 12~km from the center of the ice front, are bonded together to form a semi-circular tabular iceberg. A ring of elements whose distance, $d$, from the ice front center obeys $12 \textrm{ km} \leq   d \leq14.4 \textrm{ km}$, are allowed to move freely, but have all their bonds removed. Elements in this half annulus represent fragments of the ice shelf which calve into small pieces during the calving event.

\textcolor{red}{Should these two simulations be included?}
\subsubsection{Fixed iceberg velocity simulation}
To aid in the analysis of the ocean response to iceberg calving, another iceberg calving simulation was preformed, using a fixed iceberg velocity. This fixed velocity iceberg simulation is identical to the iceberg calving simulation (Section 3.6), except that the velocity of the calving icebergs is prescribed as $\vec{v} = = < 0, 0.01 >$ m/s. Since the iceberg speed is prescribed, this simulation is not fully coupled.

\subsubsection{Missing-Iceberg simulation}
A simulation is preformed to help analyze how the changing ice shelf geometry affects the sub-ice-shelf circulation and hydrography. In this simulation, calving iceberg (with the same geometry as the iceberg simulations above) is removed at the start of the simulation, and the simulation is spun up from rest. The simulation is run for 10 years, so that it is easily comparable to the Static ice shelf simulation (described in Section 3.5).

%\section{Results}
\section{Static ice-shelf simulation results}
\subsection{Static ice-shelf simulation}
The results in the Lagrangian static ice-shelf simulation fit within the current understanding of ice-shelf cavity circulations based on ice-shelf observations \citep{MacAyeal1984, Lewis1986, Jacobs2011} and previous modeling efforts \citep{Determan1994, Holland2006, Losch2008}.
The ocean temperatures inside the domain are warmer than the in-situ freezing point (Figure \ref{fig:static_solo_temp_v_v}a), and cause melting at the ice-shelf base (Figure  \ref{fig:static_solo_melt}a). The meltwater entering the domain is more buoyant than the water around it, and rises along the ice shelf as a cool fresh plume (Figure \ref{fig:static_solo_temp_v_v}b,c). 
%As the plume rises, it entrains ambient water causing a warming of the upper ocean (Figure \ref{fig:static_solo_temp_v_v}a).
This injection of positive buoyancy at depth drives a clockwise circulation outside of the ice-shelf cavity (Figure \ref{fig:Bt_comparison}a), providing the ice-shelf cavity with a continuous supply of warm water, which provides the thermal energy required for continuous ice-shelf melt.
Strong meridional jets are observed at the ice front along the flanks of the domain, fluxing water into the ice shelf cavity on the eastern flank and allowing water to leave the ice shelf cavity on the western side (Figure \ref{fig: meridional_velocity_snapshots}a). The meridional flow in the center of the ice front is much smaller, consistent with the dynamical barrier at the ice shelf front discussed in previous modeling and experimental studies \citep{Holland2001, Stern2014}.

%\subsubsection{Comparison with Eulerian ice-shelf model}

\subsection{Comparison of Lagrangian and Eulerian ice-shelf models}
The Lagrangian ice-shelf model results are qualitatively similar to most of the simulations from the MISOMIP experiment \citep{Asay-Davis2016}, which use a similar configuration (see Section 3.4). To get a quantitative comparison, we compare the Lagrangian ice-shelf model results to a simulation using an Eulerian ice-shelf model with an identical configuration. 
Since the Lagrangian and Eularian ice shelf models are coupled to the same ocean model and the ice shelf models use the same parametrization for ice shelf melt, we expect the results of the two models to be the almost the same, with the only differences arising from the interpolation and aggregation schemes (see Section 2.6).  This allows us to use the Eularian ice shelf model to validate the Lagrangian model.

The results show that two simulations are almost indistinguishable. This is demonstrated, for example, in Figures \ref{fig:Bt_comparison}, which shows the time-averaged barotropic stream function of the Lagrangian (Figures \ref{fig:Bt_comparison}a) and Eulerian (Figures \ref{fig:Bt_comparison}b) simulations, and the difference between the two (Figures \ref{fig:Bt_comparison}c). 
%For scale, we note that the differences between the barotropic stream functions in the two simulations, are two orders of magnitude smaller than the differences between two simulations using the same Eulerian ice shelf with different vertical coordinate systems described in Section 3.2 (not shown).
The difference between the Lagrangian and Eulerian barotropic stream functions are two orders of magnitude smaller than typical differences observed between simulations using different models in the MISOMIP \citep{Asay-Davis2016}. 
The similarity of the Lagrangian and Eulerian simulations are also reflected in the fact that the simulations have very similar ice-shelf melt rates and ocean temperature/salinity profiles (shown in Figures \ref{fig:Melt_comparison} and Figures \ref{fig:Salt_comparison} in the supplementary materials).

@@

\section{Iceberg calving simulation results}
The infrequency of large calving events together with the difficulty of placing observational instruments around tabular icebergs means that little is known about how calving icebergs affect ocean circulation around the iceberg and at the ice front. The Lagrangian ice shelf model developed in this study allows us for the first time to simulate an iceberg calving away from the ice shelf, and to study what effect this has on the ocean around the iceberg and beneath the ice shelf.  Since the Lagrangian iceberg/ice shelf model is fully coupled to the ocean, the iceberg motion affects the ocean hydrography, and the changing ocean conditions feedback onto the ice. 
%The effect of iceberg calving on the ocean around the iceberg is discussed in \citep{Stern2017}. Here we focus on how the calving event affects the region close to the ice front and beneath the ice shelf.

The results below show that the calving of a large tabular iceberg can cause significant changes to the ocean stratification and circulation around the iceberg and at the ice shelf front. 
%The changes at the ice shelf front persist, even once the tabular iceberg has drifted away from the ice shelf cavity, eventually reaching a new equilibrium consistent with the new ice front geometry. 
Changing conditions at the ice front alter the temperature fluxes into the ice shelf cavity, which impacts the hydrography inside the ice shelf cavity, and drives changes to the ice-shelf melt rates across the entire ice shelf. Local changes to the ocean hydrography are also observed around the iceberg. A Taylor column is observed below the iceberg, which travels with the iceberg as it drifts into the open ocean.

\subsection{At the ice front}
%\subsubsection{Surface warming}
After the iceberg calves, the large semi-circular tabular iceberg drifts to the northeast, driven primarily by the wind and Coriolis force. As this large iceberg and the smaller fragments moves through the water a warming of the sea surface temperature is observed around the iceberg and at the ice front (Figure \ref{fig:SST_snapshots}). The warm anomalies near the ice front persist and strengthen, even once the large iceberg fragment has drifted away. Part of this warmed surface water at the ice front is advected to west by the strong currents at the ice front (Figure \ref{fig:SST_snapshots}c). On the western side of the domain, most of this warm anomaly is advected northwards, while some of the warmth is forced southward, and is able to subduct beneath ice shelf on the western side of the ice shelf front.

%\subsubsection{Sub-surface cooling}
In addition to the surface warming, a subsurface cooling is also observed around the iceberg and at the ice front in the days following the calving event (Figure \ref{fig:Temperature_section_Collapse_yz}). 
The sub-subsurface cooling and warm surface temperature anomalies are accompanied by a freshening at depth and increase in surface salinity (not shown).
Since the background stratification has positive temperature and salinity gradients (cold/fresh water above over warm/salty), the pattern of warm salty surface anomalies above cold fresh subsurface anomalies suggests that these changes are driven by enhanced vertical mixing.

At the ice front, the subsurface negative temperature anomalies strengthen over time and within a month of the calving event, the negative anomalies occupy most of the water column close to the ice front (Figure \ref{fig:Temperature_section_Collapse_xz}). 
The zonal current along the ice front drives the cool anomaly to the west, causing an enhancement of the cool anomalies on the western side of the ice shelf front (Figure \ref{fig:Temperature_section_Collapse_xz}c).
Most of the subsurface cooling occurs outside of the ice shelf cavity, as the dynamical barrier caused by the ice shelf inhibits exchange of water across the ice front. 
However over time, an increasing amount of cooler water is able to enter the ice shelf cavity, and is advected below the ice shelf (Figure \ref{fig:Temperature_section_Collapse_yz}).

%The cooler waters entering the ice shelf cavity mix with the water within the ice shelf cavity, causing the strength of the cool anomaly to be reduced towards the grounding line (Figure \ref{fig:Temperature_section_Collapse_yz}c).

%\subsubsection{Circulation Changes}
The changing temperature flux into the ice shelf cavity in the center of the domain is partly driven by changing ocean velocities, observed at the ice front directly after the iceberg calves. The pre-calving meridional flow at the ice front is dominated by strong boundary currents moving water into and out of the ice shelf cavity along the flanks for the domain, while very little meridional velocity is observed in the center of the domain (as discussed in Section 4.1). After the iceberg calves, large meridional velocities are observed at the ice front (Figure \ref{fig:meridional_velocity_snapshots}). The structure and variability of the velocity field observed at the ice front after calving, suggest the elevated flow speed are driving by ocean instabilities that are triggered by the calving iceberg. These instabilities presumably a key driver of the enhanced vertical mixing described above. \textcolor{red}{[How do these instabilities work?]}
%The motion near the ice front remains elevated while the tabular iceberg in the vicinityand then dissipate, presumably finding a new equilibrium.

\subsection{Inside the ice shelf cavity }
%\subsubsection{Melt rate response}
The combination of cool sub-surface temperature anomalies and elevated ocean velocities at the ice front, results in a temperature flux across the ice front, as cool temperature anomalies to enter the ice shelf cavity (Figure \ref{fig:Temperature_section_Collapse_yz}). Once the cooler water has entered the ice shelf cavity, the anomalies spread southwards within the ice shelf cavity, eventually reaching all the way to the grounding line. As the cool water spreads into the cavity, it mixes with the water within the ice shelf cavity, causing a reduction in the strength of the anomaly towards the grounding line (Figure \ref{fig:Temperature_section_Collapse_yz}c).

>>>>

% and freshwater fluxes from icebergs affect regional ocean circulation, sea ice distributions and bottom water formation around Antarctica
\end{abstract}
%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%
% The body of the article must start with a \begin{article} command
%
\clearpage
\vspace{10mm}
\textbf{Key Points:}\\
\begin{itemize}
\item A novel modeling framework is developed to model breakable ice shelves and ice-shelf cavities using Lagrangian elements, held together by numerical bonds.
\item The ocean circulation beneath a (static) Lagrangian ice shelf is almost indistinguishable from the circulation beneath an Eulerian ice-shelf model run in an identical configuration, which  provides a proof of concept for the Lagrangian model. 
%\item Enhanced ocean mixing occurs near the ice front after a large calving event, which warms the ocean surface and cools the water column at depth. The cool sub-surface anomalies enter the ice shelf cavity and cause reduced basal melt in the period following the calving event. 
\item After an iceberg calves away from an ice shelf, a complex interaction is observed between the iceberg motion and ocean hydrography. This interaction leads to the formation of a Taylor column beneath the iceberg, a subsurface cooling at the ice front, and reduced melt rates within the ice shelf cavity.
\end{itemize}

@@

At the end of an ice-shelf model time step, ice-shelf fields are aggregated from the elements back onto the Eulerian ocean grid, and are then passed from the ice-shelf model to the ocean model. 
The aggregation from the Lagrangian elements onto the Eulerian grid is done by calculating the fraction of each element's volume that lies in each ocean grid cell, and dividing the fields in proportion to this fraction. For example, the amount of ice mass aggregated onto a given ocean grid cell is found by summing up the masses of all elements which intersect that grid cell, only counting the part of an element's mass that  actually intersects the ocean grid cell. When calculating the intersection between an element and a grid cell, we assume that the elements have surface areas that are shaped as regular hexagons. Seven fields are passed from the ice-shelf models to ocean model: iceberg mass and surface area (used to calculate the upper-ocean pressure field), temperature flux, salinity flux and mass flux, and meridian and zonal velocity (used to calculate the momentum flux).

@@

 \section{Experiment setup}
% the proof of concept for
%\sout{The KID model is used to represent the ice shelf in an idealized ice-shelf cavity MOM6 ocean flow. The simulations are repeated using an Eulerian ice shelf with an identical configuration. The details of the experimental setup for both the Lagrangian and Eulerian models are presented here.}
%
\subsection{Domain configuration}
In order for our simulations to be easily comparable to previous models of ice-shelf cavities, we use an experimental setup based on the configuration created for the Marine Ice Ocean Modeling Inter-comparison Project (MISOMIP) \citep{Asay-Davis2016}.
The configuration consists of an idealized ice shelf in a rectangular domain $L_{x}=80$km long and $L_{y}=480$km wide. The ice shelf is grounded on the southern side of the domain with the ice-shelf front at y=650km. The ice thickness and bottom topography of this setup are shown in Figure \ref{fig:ISOMIP_mass_and_topog}. The configuration is the same as that of the Ocean0 setup in the MISOMIP, with three changes made:
\begin{enumerate}
\item The `calving criteria' used in the MISOMIP study (which states that all points in the ice shelf with thickness less than 100m are set to zero thickness) has not been used.
\item The ice shelf has been thickened on the flanks of the domain, so that the latitude of the grounding line increases away from the center of the ice shelf.
\item The ice shelf is configured to be symmetric about its meridional center line ($x=\frac{L_{x}}{2}$). This was achieved by using the average of the left and right flanks of the ice-shelf thickness. 
\end{enumerate}
These three changes were made in order to make the circulation beneath the ice shelf easier to interpret.

\subsection{Ocean model}
The Lagrangian and Eulerian ice shelves are coupled to the MOM6 ocean model \citep{Hallberg2013}. 
The ocean model is run using a hybrid vertical coordinate system which blends a sigma-level and a z-level coordinate \citep{Stern2017}, implemented using the ALE method \citep{White2009}. In this vertical coordinate, model layers bend underneath surface topography (i.e.: the ice shelf), as they would in a sigma coordinate model, and intersect the bottom topography, as they would in a z-coordinate model. The model has 72 vertical layers and has a horizontal resolution of $\Delta x= 2$ km. The numerical simulations were all repeated using an isopycnal coordinate (without ALE regrinding-remapping). The results were qualitatively similar to the hybrid-coordinate results, and are therefore not presented here.
% All results presented below are from the ALE simulations, unless specified otherwise.

@@

%\subsection{Initializing Lagrangian elements:}
\subsection{Lagrangian ice-shelf model:}
The Lagrangian ice-shelf simulations are performed using the Kinematic Iceberg Dynamics (KID) model described in Section 2.
%\subsection{Lagrangian and Eulerian ice shelf simulations:}
The Lagrangian ice shelf is created using 10882 Lagrangian hexagonal elements with sides of length $S=0.98$ km . The positions of the hexagonal elements are initialized by packing them together in a space-filling staggered lattice. Gaps along the boundaries are filled in using smaller elements so that the total ice-shelf area is preserved. The initial mass of the ice elements is determined using bilinear interpolation from a prescribed gridded ice mass field.

\subsection{Eulerian ice-shelf model:}
The Eulerian ice-shelf simulation is performed using an existing Eulerian ice-shelf cavity model \citep{Goldberg2012}, which is an optional module of the the MOM6 ocean model. The ice shelf is initialized on the same grid as the ocean model with a horizontal resolution of $\Delta x = 2$ km. The ice-shelf thickness field is initialized using the same ice-shelf draft used for the Lagrangian model (Figure \ref{fig:ISOMIP_mass_and_topog}). The melt rates in the Eulerian ice-shelf simulation are calculated using the 3 equation model for ice-shelf decay \citep{Holland1999}, with the same parameters used to calculate the ice shelf melt rates in the Lagrangian model.

\section{Model validation}
The Lagrangian ice shelf model is validated by comparing it to an existing Eularian ice shelf model in an identical static configuration.

\subsection{Setup of static validation experiment:}
The experimental setup for the model validation experiment is based on the Ocean0 MISOMIP experiment \citep{Asay-Davis2016}.  In this experiment the ice shelves are thermodynamically active but have a time-invariant thickness. Temperature and salinity fluxes from ice shelf melt drive the circulation within the cavity.

A constant wind stress $\vec{\tau} = <\tau_{x},\tau_{y}> = < 0.05, 0.05 >$  $\frac{N}{m^{2}}$ is applied to the ocean surface. Note that this wind stress is was not applied in the original MISOMIP experiments \citep{Asay-Davis2016}.

The elements in the Lagrangian ice shelf simulation are held stationary so that the ice shelf is static and comparable to the Eularian simulation.
The models are spun up for 5 years. The analysis is performed on year 6 to 10 of the simulations.

\subsection{Results of validation experiment}

The results of the static ice shelf simulations in both simulations are qualitatively similar to results presented in \citep{Asay-Davis2016}. Melting at the base of the ice shelf drives a circulation within the cavity and strong jets are observed at the ice front and along the sides of the domain (Figure \ref{fig:Bt_comparison}). A more complete description of the circulation is provided in the supplementarity material.

Since the Lagrangian and Eularian ice shelf models are coupled to the same ocean model and the ice shelf models use the same parametrization for ice shelf melt, we expect the results of the two models to be the almost the same, with the only differences arising from the interpolation and aggregation schemes (see Section 2.6).   The results show that two simulations are almost indistinguishable. This is demonstrated, for example, in Figure \ref{fig:Bt_comparison}, which shows the time-averaged barotropic stream function of the Lagrangian (Figures \ref{fig:Bt_comparison}a) and Eulerian (Figures \ref{fig:Bt_comparison}b) simulations, and the difference between the two (Figures \ref{fig:Bt_comparison}c). 
%For scale, we note that the differences between the barotropic stream functions in the two simulations, are two orders of magnitude smaller than the differences between two simulations using the same Eulerian ice shelf with different vertical coordinate systems described in Section 3.2 (not shown).
The difference between the Lagrangian and Eulerian barotropic stream functions are two orders of magnitude smaller than typical differences observed between simulations using different models in the MISOMIP \citep{Asay-Davis2016}. 
The similarity of the Lagrangian and Eulerian simulations are also reflected in the fact that the simulations have very similar ice-shelf melt rates and ocean temperature/salinity profiles (shown in Figures \ref{fig:Melt_comparison} and Figures \ref{fig:Salt_comparison} in the supplementary materials).

@@

%\section{Results}

\section{Iceberg calving experiment}

The infrequency of large calving events together with the difficulty of placing observational instruments around tabular icebergs means that little is known about how calving icebergs affect ocean circulation around the iceberg and at the ice front. The Lagrangian ice-shelf model developed in this study allows us for the first time to simulate an iceberg calving away from the ice shelf, and to study what effect this has on the ocean around the iceberg and beneath the ice shelf.  Since the Lagrangian iceberg/ice shelf model is fully coupled to the ocean, the iceberg motion affects the ocean hydrography, and the changing ocean conditions feedback onto the ice. 
%The effect of iceberg calving on the ocean around the iceberg is discussed in \citep{Stern2017}. Here we focus on how the calving event affects the region close to the ice front and beneath the ice shelf.

In this section, the Lagrangian ice shelf model is used to simulate a tabular iceberg detaching from the ice shelf, and drifting into the open ocean. 
The results below show that the calving of a large tabular iceberg can cause significant changes to the ocean stratification and circulation around the iceberg and at the ice shelf front.

\subsection{Setup of iceberg calving experiment:}
The iceberg calving experiment is initialized using the final state of the static Lagrangian ice shelf simulation (i.e.: at time $t = 10 $ years).  When initializing the calving event, we bypass the question of how to prescribe a physical calving law \citep{Benn2007, Alley2008, Levermann2012, Bassis2013} by manually breaking off a semi-circular iceberg. 
All ice elements initially within a 14.4~km radius of the center of the ice front are allowed to move freely while the other ice elements continue to be held stationary. Ice elements less than 12~km from the center of the ice front, are bonded together to form a semi-circular tabular iceberg. A ring of elements whose distance, $d$, from the ice front center obeys $12 \textrm{ km} \leq   d \leq14.4 \textrm{ km}$, are allowed to move freely, but have all their bonds removed. 
%Elements in this half annulus represent fragments of the ice shelf which calve into small pieces during the calving event.

%\textcolor{red}{Should these two simulations be included?}
%\subsubsubsection{Fixed iceberg velocity simulation}
%To aid in the analysis of the ocean response to iceberg calving, another iceberg calving simulation was preformed, using a fixed iceberg velocity. This fixed velocity iceberg simulation is identical to the iceberg calving simulation (Section 3.6), except that the velocity of the calving icebergs is prescribed as $\vec{v} = = < 0, 0.01 >$ m/s. Since the iceberg speed is prescribed, this simulation is not fully coupled.
%
%\subsubsection{Missing-Iceberg simulation}
%A simulation is preformed to help analyze how the changing ice shelf geometry affects the sub-ice-shelf circulation and hydrography. In this simulation, calving iceberg (with the same geometry as the iceberg simulations above) is removed at the start of the simulation, and the simulation is spun up from rest. The simulation is run for 10 years, so that it is easily comparable to the Static ice shelf simulation (described in Section 3.5).

\subsection{Results of iceberg calving experiment}

%The changes at the ice shelf front persist, even once the tabular iceberg has drifted away from the ice shelf cavity, eventually reaching a new equilibrium consistent with the new ice front geometry. 
%Changing conditions at the ice front alter the temperature fluxes into the ice shelf cavity, which impacts the hydrography inside the ice shelf cavity, and drives changes to the ice-shelf melt rates across the entire ice shelf. A current is also observed circling the iceberg in a clockwise direction as the iceberg drift northwards. The  drive local changes to the ocean hydrography around the iceberg, and also influence the basin-wide circulation.
% A Taylor column is observed below the iceberg, which travels with the iceberg as it drifts into the open ocean.

%\subsection{Ocean response to iceberg detatchment}
%\subsubsection{Surface warming}
%After the iceberg calves, the large semi-circular tabular iceberg begins to drift away from the ice shelf (Figure \ref{fig:SST_snapshots}). The iceberg drifts towards the northeast, driven primarily by the wind and Coriolis force. As the iceberg drifts, it rotates in a counter-clockwise direction. After about 45 days, the iceberg collides with the eastern boundary of the domain and from their continues to drift northwards.

\subsubsection{Ocean response to iceberg detachment}
Immediately after the numerical bonds are broken, a large semi-circular tabular iceberg detaches from the ice shelf, and begins to drift northwards (Figure \ref{fig:SST_snapshots}).
The northwards motion of the iceberg creates a region of open water at the new calving front, between the ice shelf and the tabular iceberg. The formation of a region of open ocean behind the iceberg causes a sudden stretching of the water column in the wake of the iceberg, which drives an upwelling throughout the water column. The negative background temperature gradient (cold over warm) means that this upwelling is observed as warm water anomaly beneath newly formed ice front (Figure \ref{fig:Temperature_section_Collapse_yz}a and \ref{fig:Temperature_section_Collapse_xz}a). The warming is observed throughout the water column, but is largest near the surface.  Similarly, the upwelling leads to increased salinity beneath newly formed ice front (not shown).

The calving event also drives an immediate change in circulation around the newly created ice front as the ocean adjusts to the topographic changes. 
Since the Coriolis parameter is negative, a stretching of the water column behind the iceberg has to be accompanied by the creation of negative relative vorticity in order for potential vorticity (PV) to be conserved. 
The creation of negative relative vorticity along the newly calved ice front, gives rise to a pair of oppositely orientated ocean jets running along the ice front. A westward barotropic jet is created to the south of the calving front, and an eastward barotropic jet is created to the north of the calving front (Figures \ref{fig:u_velocity_horizontal}a and \ref{fig:v_velocity_vertical}b). The positive gradient in zonal velocity created by this pair of oppositely orientated jets yields the negative relative vorticity, needed to conserve PV.

To the north of the iceberg, the iceberg detachment has the opposite effect on the ocean. The forward movement of the iceberg causes a squeezing of the water column directly in front of the iceberg. This shortening of the water column causes a downwelling in the water column in front of the iceberg, which is observed as a cool water anomaly in front of the iceberg (Figure \ref{fig:Temperature_section_Collapse_yz}a). The squeezing of the water column generates positive relative vorticity, so that the total PV is conserved. The positive relative vorticity creates an eastward and westward jet to the south and north of the northern edge of the iceberg (Figures \ref{fig:u_velocity_horizontal}a and \ref{fig:v_velocity_vertical}b). The stretching and squeezing of the water column to the south and north of the iceberg, both contribute to the eastward jet which forms directly beneath the iceberg which plays an important role in driving iceberg motion.
%The eastward jets created beneath the iceberg merge into one combined jet, so that the flow beneath the iceberg is towards the east.

 \textcolor{red}{(If you have time, check whether when the wind goes in the other direction, the jet beneath the iceberg moves in the other direction.)}
%\begin{equation}
%q = \frac{f + \zeta}{h}
%\end{equation}
\subsubsection{Iceberg motion}
After the iceberg breaks away from the ice shelf, the motion of iceberg is primarily driven by the drag from ocean currents beneath the iceberg, wind drag and by Coriolis force. The eastward jet beneath the iceberg drives the iceberg towards the east (Figures \ref{fig:u_velocity_horizontal}a and \ref{fig:v_velocity_vertical}b), while the wind and Coriolis force drive the iceberg offshore. Initially, as the iceberg drifts eastward, it collides with the ice shelf, which hinders its motion (Figure \ref{fig:SST_snapshots}a). Only once the iceberg has drifted sufficiently far northwards to be clear of the ice shelf, does it begin to drift towards the east. Once this happens, the iceberg velocity quickly adjusts so that it is approximately equal to the ocean velocity below the iceberg. % \textcolor{red}{(check this is true)}.
As the iceberg drifts northwards, it rotates in a counter-clockwise direction (Figure \ref{fig:SST_snapshots}).

%The rotation is partly due to the semi-circular shape of the iceberg being guided by the semi-circular ice front, but is also observed in experiments using rectangular icebergs, where it is mainly caused by the Coriolis force. The iceberg continues to rotate even after it has moved entirely away from the ice shelf. 
The motion of the iceberg plays a large role in setting the direction of the ocean flow below and around the iceberg, since the ocean has to adjust to topographical changes so that PV is conserved.  This is illustrated, for example, by the change in orientation of the ocean jets running beneath the iceberg that occurs as the iceberg rotates (Figure \ref{fig:u_velocity_horizontal}b). An important consequence of the controlling effect of iceberg topography is formation of a Taylor Column below the iceberg: the water column below the iceberg is constrained to move at the same speed as the iceberg above, since differential motion would force the water column out from under the iceberg, causing the water column to stretch, and generating PV.  This means that the column of water beneath the iceberg is largely separated from the water around the iceberg and strong gradients in temperature, salinity and velocity can exist at the iceberg edge. In our experiment, a Taylor Column is clearly observed below the iceberg throughout the simulation (Figure \ref{fig:Taylor_column}). %(the changes in PV associated with the rotation of the jets are ultimately driven by the external force of the wind)

Once the iceberg reaches the eastern side of the domain, it begins to interact the boundary of the domain and with the southward boundary current running along the edge of the domain (Figures \ref{fig:SST_snapshots}b and \ref{fig:v_velocity_horizontal}c). 
The presence of the iceberg along the eastern boundary impedes the southward boundary current, so that the current has to flow to the west of the iceberg (Figure \ref{fig:v_velocity_horizontal}b). 
The diverted current follows the topographic contour created by the iceberg so that it sets up a  counter-clockwise circulation running around the iceberg (Figure \ref{fig:u_velocity_horizontal}c). The eastward jet which was previously positioned beneath the iceberg shifts to the south and becomes part of the flow directing the boundary current around the iceberg. This counter-clockwise circulation around the iceberg remains around the iceberg until the end of the simulation. Interestingly, the iceberg continues to move northward despite the fact that it is moving within a southward flowing boundary current (Figure \ref{fig:Taylor_column}b). This is partly due to the Taylor Column beneath the iceberg, which allows the northward velocity beneath the iceberg to be separated from the southward moving current around the iceberg.

%The iceberg continues to rotate as it approaches the side of the domain, while the mean iceberg motion is directed to move parallel to the eastern boundary so that the iceberg motion is towards the north.

%The diverting of the flow towards the west is aided by the westward current in front of the iceberg (discussed above).

%\textcolor{red}{Mention: Similar Taylor columns have been observed above sea mounts [reference].}

%\textcolor{red}{Possibly include: Directly below the iceberg on the ocean floor, an increased dipycnal mixing is observed in the bottom boundary layer. This elevated mixing in the bottom boundary layer is likely a result of a barotropic adjustment driven by the iceberg motion. As the iceberg drifts through the water, the region of elevated dipycnal mixing shifts so that it remains directly below the iceberg. Similarly elevated rates in dipycnal mixing are observed at the bottom of the water column near the ice front. }
